### 2015-03-23 markdown/readme.tex
towards kramdown integration

\section{MS Prep for Genetic Programming Theory \& Practice 2015}\hypertarget{ms-prep-for-genetic-programming-theory--practice-2015}{}\label{ms-prep-for-genetic-programming-theory--practice-2015}

\subsection{In brief}\hypertarget{in-brief}{}\label{in-brief}

``Genetic programming as if you meant it: real and imaginary user experience''

My title is a direct nod to \href{http://cumulative-hypotheses.org/2011/08/30/tdd-as-if-you-meant-it/}{Keith Braithwaite's} exercise for software developers, ``TDD as if you meant it''. Like Keith's exercise, what I describe here for genetic programming will seem onerous and difficult to most experienced theoreticians or practitioners. I'll make the case that this approach suggests---and demonstrates---a broader systematic framework in which genetic programming is far more likely to \emph{provide value to its users} than we find in a quarter-decade of exploration. Along the way, I'm forced to address deep philosophical problems in the program of genetic progamming research and practice more broadly: \emph{viz.}, what it is ``for'' and when it is ``done''.