### 2015-03-27 tex/Book/gpbook-Tozier-2015-03-23/chapter.tex
second pass, first part

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title*{GP As If You Meant It: Real and Imaginary User Experience}
\titlerunning{GP As If You Meant It}

% your contribution title if the original one is too long
\author{William A. Tozier}
\authorrunning{Tozier}

%\institute{William A. Tozier}

\maketitle

\abstract{(to be written last)}

\begin{keywords}
keywords to your chapter, these words should also be indexed
\end{keywords}
\index{keywords to your chapter}
\index{these words should also be indexed}
\\

\subimport{../../../markdown/}{manuscript}

\bibliographystyle{spbasic}
\bibliography{gp-bibliography,chapter}