### 2015-03-28 markdown/session.tex

\section{An example session}\hypertarget{an-example-session}{}\label{an-example-session}

(test placeholder)

\begin{description}
\item[{\tt me}] I do and say a thing here

This is a continuation of the first item, and should include a few extra fillips, like $x=y$ and a little

\begin{itemize}
\item list
\item of
\item items
\end{itemize}

And also it includes a

\begin{quote}
blockquote right here
\end{quote}

\item[{\tt system}] it does and says something back to me

\item[{\tt me}] he look I can have a little dialog sortof

\item[{\tt system}] I can fight you every step of the way, though!
\end{description}