### 2015-03-28 markdown/session.tex
scenario placeholder'

\section{An example session}\hypertarget{an-example-session}{}\label{an-example-session}

\textbf{Placeholder}: I am converting the transcript of the session from an awful multipage LaTeX {\tt \textbackslash{}longtable} into a much more manageable {\tt \textbackslash{}enum} list like this. It will take a while, since the original {\tt \textbackslash{}longtable} was more than nine pages long and full of table markup.

\subsection{The stage is set}\hypertarget{the-stage-is-set}{}\label{the-stage-is-set}

\begin{description}
\item[{\tt me}] well what have we here?

\item[{\tt system}] it does and says something back to me

\item[{\tt me}] hey look I can have a little dialog sortof

\item[{\tt system}] I can fight you every step of the way, though!
\end{description}

\subsection{Progress and Stagnation}\hypertarget{progress-and-stagnation}{}\label{progress-and-stagnation}

\begin{description}
\item[{\tt me}] Keep going

\item[{\tt system}] OK!
\end{description}

\subsection{Innovation}\hypertarget{innovation}{}\label{innovation}

\begin{description}
\item[{\tt me}] Hey!

\item[{\tt system}] What?
\end{description}

\subsection{Frustrations}\hypertarget{frustrations}{}\label{frustrations}

(\&c)